Equalizer Programming Guide

### 2009-03-10 Developer/ProgrammingGuide/paper.tex
Update Programming Guide with new AGL event dispatch
[ ] May break build
[ ] Breaks existing applications
[ ] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[x] Documentation

\date{
  \vspace{15cm}
  {\Large\sffamily The official reference for creating parallel,
    scalable OpenGL\texttrademark\ applications with the Equalizer
    parallel rendering framework}\\\vspace{1cm}
  Version 1.5.4 for Equalizer 0.6+svn\\[\medskipamount]
%  \today
}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
\listoffigures
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.5.1   & Dec 08, 2008 & Minor updates based on the print version\\
    1.5.2   & Jan 10, 2009 & Added \sref{sCompositor}\\
    1.5.3   & Jan 13, 2009 & Added \sref{sNetwork}\\
    1.5.4   & Feb 09, 2009 & Added \sref{sNetUsage}\\
    \hline
  \end{tabularx}}
\clearpage

@@

In general, events are received and processed by the pipe thread a
window belongs to. An exception to this rule is AGL, where all events
are dispatched from the main node thread. WGL and GLX receive and
process the events from the pipe threads that created the
windows. Whenever the term \textbf{event thread} is used, it refers to
the thread receiving the event, i.e., the pipe thread for WGL and GLX,
and the main thread for AGL.

@@

To dispatch events, Equalizer 'pumps' the native events. On WGL and GLX,
this happens on each thread with windows, whereas on AGL it happens only
on the main thread. By default, Equalizer pumps these events
automatically for the application in-between executing task methods.

@@

The method \textsf{useMessagePump} is called by Equalizer during
application and pipe thread initialization (before configInit!) to
determine if the pipe's thread shall automatically dispatch OS
events. For non-threaded and AGL pipes, the method affects the node
thread. On WGL and GLX, messages are dispatched in the application and
each pipe thread. On AGL, events are dispatched only on the application
thread.

@@

Carbon events issued for AGL windows are processed from the node's main
thread. One \textsf{AGLEventHandler} per process is used. When a Carbon
window is set on an \textsf{AGLWindow}, the window is registered with
the global \textsf{AGLEventHandler}. The event handler installs a Carbon
event handler for all important events. The event handler uses an
\textsf{AGLWindowEvent} to pass the Carbon \textsf{EventRef} to
\textsf{AGLWindowIF::processEvent}. During window exit, the installed
Carbon handler is removed when the window is deregistered with the
\textsf{AGLEventHandler}. No event handling is set up for PBuffers.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\small\label{fSlices}Volume Rendering using 3D Textures}}
\end{wrapfigure}
When using 2D (sort-first) or stereo decompositions, no special
programming is needed to achieve good scalability, as \textsf{eVolve} is
mostly fill-limited and therefore scales nicely in these modes.

The full power of scalable volume rendering is however in DB (sort-last)
compounds, where the full volume is divided into separate bricks. Each
of the bricks is rendered like a separate volume. For recomposition, the
\textsf{RGBA} frame buffer data resulting from these render passes then
has to be assembled correctly.

@@

For recomposition, the 2D frame buffer contents are blended to form a
seamless picture. For correct blending, the frames are ordered in the
same back-to-front order as the slices used for rendering, and use the
same blending parameters. Simplified, the frame buffer images are
`thick' slices which are `rendered' by writing their content to the
destination frame buffer using the correct order.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/volResult.png}
  {\caption{\small\label{fVolResult}Result of \fig{fBlendPerspective}}}
\end{wrapfigure}
For orthographic rendering, determining the compositing order of the
input frames is trivial. The screen-space orientation of the volume
bricks determines the order in which they have to be composited. The
bricks in \textsf{eVolve} are created by slicing the volume along one
dimension. Therefore the range of the resulting frame buffer images,
together with the sorting order, is used to arrange the frames during
compositing. \fig{fBlendOrtho} shows this composition for one view.

@@

Distributed objects provide powerful, object-oriented data
distribution for C++ objects. Their functionality and an example use
case for parallel rendering has been described in \sref{sNetObject}.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/netObject.pdf}
  {\caption{\small\label{fNetObject}Communication between two objects}}
\end{wrapfigure}
Multiple instances of the same object (identifier) can be mapped on the
same node and session simultaneously. Each object instance has a
node-unique instance identifier. When the instance ID is set to
\textsf{EQ\_ID\_ANY} (the default value), a command is dispatch to all
object instance registered to the the session with the correct object
ID. Multiple object instance per node are used when different threads on
the node need to process a different version of the same object.

>>>>

\date{
  \vspace{15cm}
  {\Large\sffamily The official reference for creating parallel,
    scalable OpenGL\texttrademark\ applications with the Equalizer
    parallel rendering framework}\\\vspace{1cm}
  Version 1.5.5 for Equalizer 0.6+svn\\[\medskipamount]
%  \today
}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
\listoffigures
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.5.1   & Dec 08, 2008 & Minor updates based on the print version\\
    1.5.2   & Jan 10, 2009 & Added \sref{sCompositor}\\
    1.5.3   & Jan 13, 2009 & Added \sref{sNetwork}\\
    1.5.4   & Feb 09, 2009 & Added \sref{sNetUsage}\\
    1.5.5   & Mar 10, 2009 & Updated to reflect new AGL event dispatch\\
    \hline
  \end{tabularx}}
\clearpage

@@

Events are received and processed by the pipe thread a window belongs
to. For AGL, Equalizer internally forwards the event from the main
thread, where it is received, to the appropriate pipe thread.

@@

To dispatch events, Equalizer 'pumps' the native events. On WGL and GLX,
this happens on each thread with windows, whereas on AGL it happens on
the main thread and each pipe thread. By default, Equalizer pumps these
events automatically for the application in-between executing task
methods.

@@

The method \textsf{useMessagePump} is called by Equalizer during
application and pipe thread initialization (before configInit!) to
determine if the pipe's thread shall automatically dispatch OS
events. For AGL, this affects the node threads and pipe thread, since
the node thread message pump needs to dispatch the events to the pipe
thread event queue.

@@

Carbon events issued for AGL windows are initially receive by the node
main thread, and automatically dispatched to the pipe thread. The pipe
thread will dispatch the event to the appropriate event handler. One
\textsf{AGLEventHandler} per window is used, which is created during
\textsf{AGLWindow::configInit}. The event handler installs a Carbon
event handler for all important events. The event handler uses an
\textsf{AGLWindowEvent} to pass the Carbon \textsf{EventRef} to
\textsf{AGLWindowIF::processEvent}. During window exit, the installed
Carbon handler is removed when the window's event handler is deleted. No
event handling is set up for PBuffers.

@@

When using 2D (sort-first) or stereo decompositions, no special
programming is needed to achieve good scalability, as \textsf{eVolve} is
mostly fill-limited and therefore scales nicely in these modes.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\small\label{fSlices}Volume Rendering using 3D Textures}}
\end{wrapfigure}
The full power of scalable volume rendering is however in DB (sort-last)
compounds, where the full volume is divided into separate bricks. Each
of the bricks is rendered like a separate volume. For recomposition, the
\textsf{RGBA} frame buffer data resulting from these render passes then
has to be assembled correctly.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \vspace{-3ex}\includegraphics[width=.382\textwidth]{images/volResult.png}
  {\caption{\small\label{fVolResult}Result of \fig{fBlendPerspective}}}
  \vspace{-5ex}
\end{wrapfigure}
For recomposition, the 2D frame buffer contents are blended to form a
seamless picture. For correct blending, the frames are ordered in the
same back-to-front order as the slices used for rendering, and use the
same blending parameters. Simplified, the frame buffer images are
`thick' slices which are `rendered' by writing their content to the
destination frame buffer using the correct order.

For orthographic rendering, determining the compositing order of the
input frames is trivial. The screen-space orientation of the volume
bricks determines the order in which they have to be composited. The
bricks in \textsf{eVolve} are created by slicing the volume along one
dimension. Therefore the range of the resulting frame buffer images,
together with the sorting order, is used to arrange the frames during
compositing. \fig{fBlendOrtho} shows this composition for one view.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/netObject.pdf}
  {\caption{\small\label{fNetObject}Communication between two objects}}
\end{wrapfigure}
Distributed objects provide powerful, object-oriented data
distribution for C++ objects. Their functionality and an example use
case for parallel rendering has been described in \sref{sNetObject}.

@@

Multiple instances of the same object (identifier) can be mapped on the
same node and session simultaneously. Each object instance has a
node-unique instance identifier. When the instance ID is set to
\textsf{EQ\_ID\_ANY} (the default value), a command is dispatch to all
object instance registered to the the session with the correct object
ID. Multiple object instance per node are used when different threads on
the node need to process a different version of the same object.
